---
title: "CPUE and L90 calculations"
author: "Matilda Andersson"
date: "2024-06-18"
output: html_document
---

load prackages
```{r}
library(tidyverse)
```

load data
```{r}
Catch_20232019 <- read_csv("Catch_20232019.csv")
```

calculate effort (number of nets)
You could add net depth if you want to be able to look at CPUE at diff depths or remove one depth from the analysis
```{r}
Effort_Nnets <- Catch_20232019 %>%
  group_by(Year, Location) %>%
  summarise(Effort_Nnets = n_distinct(Station), .groups = 'drop')
```

remove individuals <12 cm -- not caught representativly in the monitoring nets
```{r}
Catch_20232019 <- Catch_20232019 %>%
  filter(Length_cm >= 12)
```

group-by year:location:species -- count individuals (there is one row per individual)
```{r}
Indiv_count <- Catch_20232019 %>%
  group_by(Year, Location, Species) %>%
  summarise(count = n(),
    Latitude = first(Latitude),
    Longitude = first(Longitude),
    .groups = 'drop')
```

calculate CPUE
```{r}
#join the count and effort dataframes
CPUE_data <- Indiv_count %>%
  left_join(Effort_Nnets, by = c("Year", "Location"))

CPUE_results <- CPUE_data %>%
  mutate(CPUE = count / Effort_Nnets)

write.csv(CPUE_results, "CPUE_results_20232019.csv", row.names = FALSE)
```

calculate L90
if statements 
if Abbore -- length >= 15
if other use the general filter of >= 12
if # of individuals < 50 do not calculate -- error message -- "Your numbers look a bit fishy -- locations have been removed due to low sample size"
```{r}
L90_data <- Catch_20232019 %>%
  filter(!(Species == "Abborre" & Length_cm < 15)) %>%
  filter(!(Species != "Abborre" & Length_cm < 12))

L90 <- L90_data %>%   
  group_by(Year, Location, Species) %>%
  summarise(median_cm=quantile(Length_cm, probs=0.5),
            L90_cm=quantile(Length_cm, probs=0.9),
            max_cm=max(Length_cm),
            Ntot=length(Length_cm))

L90 <- L90 %>%
  filter((Ntot >= 50))

write.csv(L90, "L90_20232019.csv", row.names = FALSE)
```




