---
title: "DataCleaning2"
author: "Matilda Andersson"
date: "2024-06-15"
output: html_document
---
Code to clean and reorganize the fish data to make it usable in a shiny app


load packages
```{r}
library(tidyverse)
```

load data
```{r}
Catch_raw_20232019 <- read_csv("data/Detaljuttag då längd inkl tomfångst.csv", col_types = cols(
  År1 = col_double(),   # Specify 'species' as character
  Lokal = col_character(),   # Specify 'species' as character
  Fångstområde = col_character(),
  Djupstratum1 = col_character(),
  störning1 = col_character(),
  Artbestämning1 = col_character(),
  Längdgrupp = col_double(),
  Antal1 = col_double(),
  Vikt_uppmätt = col_character(),
  Lat_grader = col_character(),
  Long_grader = col_character(),
))

Catch_raw_20232019 <- as_tibble(Catch_raw_20232019)
```

selected required columns and change from , to . as decimal point
```{r}
Catch_20232019 <- Catch_raw_20232019 |> select(År1, Lokal, Fångstområde, Djupstratum1, störning1, Artbestämning1, Längdgrupp, Antal1, Vikt_uppmätt, Lat_grader, Long_grader)

clean_up_commas <- function(x) {
  x <- gsub(",", ".", x)  # Replace commas with periods
}

Catch_20232019 <- Catch_20232019 |> 
  mutate(
    Vikt_uppmätt = clean_up_commas(Vikt_uppmätt),
    Djupstratum1 = clean_up_commas(Djupstratum1),
    Lat_grader = clean_up_commas(Lat_grader),
    Long_grader = clean_up_commas(Long_grader))
```

År1 - Fishing year
Lokal - Region (More general location)
Fångstområde - Location (More specific location)
Djupstratum1 - Net depth
störning1 - Was the net disturbed (only want no)
Artbestämning1 - Species
Längdgrupp - Length (cm)
Antal1 - Number of fish
Vikt_uppmätt - Weight (kg)
Lat_grader - Lat
Long_grader - Long


re-name columns 
```{r}
Catch_20232019 <- Catch_20232019 |> 
  rename(Year = År1,
         Region = Lokal,
         Location = Fångstområde,
         Net_depth = Djupstratum1,
         Disturbance = störning1,
         Species = Artbestämning1,
         Length_cm = Längdgrupp,
         Number = Antal1,
         Weight =  Vikt_uppmätt,
         Latitude = Lat_grader,
         Longitude = Long_grader)
```

How many species are there - how abundant - decide what to remove
```{r}
Catch_20232019 %>%
  count(Species)
```

Remove the species that are likely a mistake or were caught fewer than 10 times in the past 5 years
```{r}
Catch_20232019 <- Catch_20232019 %>%
  filter(!Species %in% c("Ullhandskrabba", "Staksill", "Lax/öring obestämd", "Ejder", "Ål", "Tunga", "Småspigg"))
```

re-name species -- maybe -- but we can come back to this
```{r}

```

group-by year:location:species:depth
```{r}

```

calculate CPUE
```{r}

```

calculate L90
```{r}

```



